Mathematical Aspects of the Capital Market

### 2017-08-10 phd/Chapterintroduction.tex
Revised the introduction chapter. reimplemented the sampling algorithm using resampling.

\section{Regular Variation}
The concept of regular variation is defined by the following scaling
property: if a function $f$ on $(0, \infty)$ satisfies
\[
\lim_{x \to \infty} {f(c x) \over f(x)} = c^\alpha
\quad
\forall c > 0
\]
then we say $f$ is regularly varying with index $\alpha$.
$f$ can be written in the form $f(x) = {L(x) \over x^\alpha}$, where
$L(x)$ is a slowly varying function, i.e.
$\lim_{x \to \infty} {L(c x) \over L(x)} = 1$, $\forall c > 0$.
We call a random variable $X$ regularly varying with index
$\alpha \geq 0$ if it satisfies the tail balance condition in the
limit $x \to \infty$.
\[
\P(X > x) \sim p_+ {L(x) \over x^\alpha},\quad
\P(X < -x) \sim p_- {L(x) \over x^\alpha}, \quad
\text{for some } p_+, p_- > 0,\; p_+ + p_- = 1
\]

@@

For an extensive summary of the regular variation properties of
functions of regularly varying random variables, see Mikosch and
Jessen \cite{JessenMikosch2006}.

@@

Later authors, e.g. Back \cite{back:1991} chose to model the log-price
process as a semimartingale, with increments of the martingale
component modelled as a product process:
\begin{eqnarray*}
  Y_t &=& Y_0 + A_t + M_t \\
  M_t - M_{t-1} &=& X_t = \sigma_t Z_t
\end{eqnarray*}
where $A_t$ is a finite variation process and $M_t$ is a martingale
and hence $Y_t$ is a semimartingale. $\sigma_t$ is non-negative and
$Z_t$ is an iid process with 0 mean and unit variance. A convenient
choice of $\sigma_t$ is
\[
\log\sigma_t = \sum_{l \in \mathbb Z} \psi_l \eta_{t-l}
\]
where $-1 \leq \{\psi_l\}_{l \in \mathbb Z} \leq 1$
is a sequence of real numbers with at least one non-zero element.

>>>>

\section{Regular Variation}\label{sec:intro_regvar}
The concept of regular variation is defined by the following scaling
property: if a function $f$ on $(0, \infty)$ satisfies
\[
\lim_{x \to \infty} {f(c x) \over f(x)} = c^\alpha
\quad
\forall c > 0
\]
then we say $f$ is regularly varying with index $\alpha$.
$f$ can be written in the form $f(x) = {L(x) \over x^\alpha}$, where
$L(x)$ is a slowly varying function, i.e.
$\lim_{x \to \infty} {L(c x) \over L(x)} = 1$, $\forall c > 0$.
We call a random variable $X$ regularly varying with index
$\alpha \geq 0$ if it satisfies the tail balance condition in the
limit $x \to \infty$.
\[
\P(X > x) \sim p_+ {L(x) \over x^\alpha},\quad
\P(X < -x) \sim p_- {L(x) \over x^\alpha}, \quad
\text{for some } p_+, p_- > 0,\; p_+ + p_- = 1
\]

@@

In addition to Breiman's result, the following is also well-known,
assuming $X_1, X_2$ are positive independent random variables and
$X_1$ is regularly varying with index $\alpha$:
\begin{enumerate}
\item if $X_2$ is regularly varying with index $\alpha$ or
  $\lim_{x \to \infty} {P(X_2 > x) \over \P(X_1 > x)} = 0$, then $X_1
  X_2$ is regularly varying with index $\alpha$.
\item if $X_1, X_2$ are iid and $\E X_1^\alpha = \infty$, then
  ${\P(X_1 X_2 > x) \over \P(X_1 > x)} \to \infty$
\item if $X_1, X_2$ are iid and $\E X_1^\alpha < \infty$, then the
  only possible limit of ${\P(X_1 X_2 > x) \over \P(X_1 > x)}$ is
  $2 \E X_1^\alpha$.
\end{enumerate}
For an extensive summary of the regular variation properties of
functions of regularly varying random variables, see Mikosch and
Jessen \cite{JessenMikosch2006}.

@@

Later authors, e.g. Back \cite{back:1991} chose to model the log-price
process as a semimartingale, with increments of the martingale
component modelled as a product process:
\begin{eqnarray*}
  Y_t &=& Y_0 + A_t + M_t \\
  M_t - M_{t-1} &=& X_t = \sigma_t Z_t
\end{eqnarray*}
where $A_t$ is a finite variation process and $M_t$ is a martingale
and hence $Y_t$ is a semimartingale. $\sigma_t$ is non-negative and
$Z_t$ is an iid process with 0 mean and unit variance. A convenient
choice of $\sigma_t$ is
\begin{equation}
  \label{eq:rfht}
  \log\sigma_t = \sum_{l \in \mathbb Z} \psi_l \eta_{t-l}  
\end{equation}
where $-1 \leq \{\psi_l\}_{l \in \mathbb Z} \leq 1$
is a sequence of real numbers with at least one non-zero element,
$\{\eta_t\}$ is a iid sequence of random variables. By Kolmogorov's
3-series theorem, the infinite series above converges if and only if
$\sum_{l \in \mathbb Z} \psi_l^2 < \infty$.
In particular, if $\{\eta_i\}$ is normally distributed with mean 0,
$\{\sigma_t\}$ is stationary and its marginal distributions are
log-normal.

If, however, the sequence $\{Z_t\}$ or $\{\sigma_t\}$ is regularly
varying with index $\alpha$ and some additional conditions are
satisfied, $X_t$ is also regularly varying with the same
index. Specifically, if $Z_t$ is regularly varying and $\sigma_t$ has
a lighter tail, the conclusion follows from Breiman's lemma. See
\S\ref{sec:case1} of Jan\ss en et al. i.e. chapter \ref{ch:bernoulli}
of this thesis for more details.

If, instead, $\sigma_t$ has a heavier tail than $Z_t$, $X_{i,t}$ is
regularly varying assuming the following conditions:
\begin{itemize}
\item $e^\eta$ is regularly varying with index $\alpha$, where $\eta$
  is a generic element of the iid sequence $\{\eta_t\}$, and
  \begin{equation}
    \label{eq:dfhty}
    \E e^{\alpha \eta} = \infty, \quad\text{ or }
    {\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)} \to c \in (0, \infty)
  \end{equation}
    where $\eta, \eta_1$ and $\eta_2$ are generic elements of the
    iid sequence $\{\eta_t\}$.
\end{itemize}
Note the only finite limit for
${\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)}$ in $(0, \infty)$ is
$\E e^{\alpha \eta}$. Hence the two conditions of \eqref{eq:dfhty}
are mutually exclusive. In chapter \ref{ch:bernoulli} we will see that
these two conditions lead to rather different behavior for
eigenvalues of the sample covariance matrix of stochastic volatility
processes of the form \eqref{eq:rfht}.

There are a few advantages in using stochastic volatility models. For
one thing, they are among the simplest models allowing for conditional
heteroscedasticity; for another thing,
